Towards a general mathematical theory of experimental science

### 2019-02-08 Articles/TowardsAGeneralMathematicalTheoryOfExperimentalScience/article.tex
Reviewed introduction and conclusions

\documentclass[letterpaper]{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[letterpaper,top=1in,bottom=1in,left=1in,right=1in,marginparwidth=0.5in]{geometry}

% Temporary to remove "submitted to"
\makeatletter
\def\ps@pprintTitle{%
	\let\@oddhead\@empty
	\let\@evenhead\@empty
	\def\@oddfoot{}%
	\let\@evenfoot\@oddfoot}
\makeatother

\usepackage{lineno,hyperref}

\usepackage{amsmath, amsthm, amsfonts}
\usepackage[only,llbracket, rrbracket,llparenthesis,rrparenthesis]{stmaryrd}

@@

\begin{abstract}
	In this article we lay the groundwork for a general mathematical theory of experimental science. The starting point will be the notion of verifiable statements, those assertions that can be shown to be true with an experimental test. We study the algebra of such objects and show how it is closed only under finite conjunction and countable disjunction. With simple constructions, we show that the set of possible cases distinguishable by verifiable statements is equipped with a natural Kolmogorov and second countable topology and a natural $\sigma$-algebra. This gives a clear physical meaning to those mathematical structures and provides a strong justification for their use in science. It is our hope and belief that such an approach can be extended to many areas of fundamental physics and will provide a consistent vocabulary across scientific domains.
\end{abstract}

% We form experimental domains, which represent the list of all possible testable answers to a particular scientific question. From these we construct theoretical domains, which represent all possible predictions (not necessarily verifiable themselves). We then define the possibilities of an experimental domain as those predictions that, if true, predict the truthfulness of all verifiable statements in the domain. We then show that experimental domains provide a natural Kolmogorov and second countable topology over their possibilities, while theoretical domains provide a natural $\sigma$-algebra.

@@

During the course of the last century a few fields of science and engineering have developed an associated general theory. For example, in computer science theory one starts by defining what a computational device is~\cite{Turing} and then shows that no algorithm exists that can correctly decide whether an arbitrary program terminates given an arbitrary input~\cite{Sipser}. In communications one defines what information is~\cite{Shannon} and shows that no encoding can outperform the Huffman coding~\cite{Pierce}. In control theory one defines what a system is~\cite{Brogan} and then looks for general strategies for control such as Kalman filtering~\cite{Kalman}. These theories are general in the sense that their results respectively apply to all control systems, all communication systems and all computational devices. That is, the results apply because of the mere definition of the subject matter.

No such general theory currently exists for physics, which is essentially composed of different theories loosely connected to each other, each with their own realm of applicability. However, we believe such a general theory can be created and in this work we present our basic strategy towards it. As one familiar with the general theories of other subjects would expect, we only need to focus on clarifying exactly what our starting points are and let the logic follow. That is, once we formalize exactly what it means to study a system through experimentation, we can reach interesting results that necessarily apply to all scientific theories. For example, we will see that any set of objects that are physically distinguishable must have cardinality no greater than continuum. We will see that defining what can be experimentally verified about them will always induce a topology that is Kolmogorov and second countable. These results are general because they apply independently of what is being studied.

One should not confuse this approach with the idea of a grand-unified theory or theory of everything. In that program, one looks for the theory of fundamental constituents from which all other theories can be derived. Our program is altogether different and orthogonal. Our subject matter is the scientific models themselves. Our aim is to understand what aspects of a scientific theory are already constrained by the mere fact that it is a scientific theory or by the assumptions one builds into them. While this general theory may be of some utility to someone working on a grand-unified theory as it may help narrow the search, neither one will lead to the other.

In this paper we briefly present the starting point for our general mathematical theory of experimental science. It is the product of a decade spent in reverse engineering the laws of particle mechanics~\cite{Carc1} and our subsequent work on formalizing the starting points~\cite{Carc2}. Our overall program is large and ambitious and a journal article cannot treat it comprehensively. Therefore we will only present an overview of the basic concepts, which are already well consolidated and provide a good template to understand our approach. We will only provide a sketch for the proofs, which can be found in the draft of our open access book~\cite{Carc3} together with more details, discussions and insights.

The basic idea is that science deals with verifiable statements: assertions whose truth can be tested experimentally. We study the logic of these statements, which is different from the standard Boolean logic. We then group them into experimental domains, which represent a set of verifiable statements that can be tested in an indefinite amount of time. From these we construct theoretical domains, which represent all possible predictions, and within these we find the possibilities, those predictions that determine the outcome of all possible tests. We will find that possibilities have a natural topology induced by the experimental domain and a natural $\sigma$-algebra induced by the theoretical domain. We will find that the topology is Kolmogorov and second countable, which limits the cardinality of the set of possibilities of a domain to the one of continuum. Note that these structures are indeed pervasive in science, and form the foundation for differential geometry (Riemannian and symplectic), measure theory, probability theory and many other mathematical tools already used in physics and other sciences.

@@

The first thing we need to do is to characterize what science is and what its scope is. We do so by introducing the following:

@@

This means that science restricts itself to the study of assertions that have a well defined truth value which can be verified experimentally. The issue at hand is to formally capture this informal intuition. We start with the following common definition.

@@

Next we need to define our truth bearer. In mathematical logic, this is typically a well formed formula. This cannot work for us: what we are interested in is the meaning of the assertion and not how it is expressed. For example, \statement{this animal is a dog} and \statement{questo animale \`e un cane} represent the same fact expressed in different languages. As science is universal, it should not matter the language, units or reference system used to make an assertion. While we will still use standard mathematical logic for the formal system, we introduce a variation of algebraic logic to represent our ``informal'' assertions. We start by defining our truth bearer as:

@@

\section{Brief discussion}

The connection we outlined creates a strong bridge between the standard mathematical structures used in physics and their meaning. Every theorem, every proof on those structures can now be given a direct physical meaning as well. And given the foundational nature of topologies and $\sigma$-algebras, we hope to extend this framework to measure theory, differential geometry, symplectic geometry, Riemannian geometry, probability theory and so on.

To give an example, consider an experimental domain where the basis is composed of statements like \statement{this quantity is between $a$ and $b$} where $a$ and $b$ are different values that can be arbitrarily close. For example, \statement{the distance between the earth and the moon is between 384 and 385 thousand Km}. Each verifiable statement of the basis corresponds to an open interval of the real line, therefore we find a correspondence between arbitrary precision measurements and the standard topology on the reals, as this is the one generated by open intervals. Note, in fact, that this topology is second-countable and Kolmogorov.

The possibilities (i.e. \statement{the mass of the photon is precisely zero}) are not themselves verifiable since infinite precision measurements of a continuous quantity are not possible. Yet, their negation (i.e. \statement{the mass of the photon isn't precisely zero}) could be verified in practice. This is expressed mathematically by the fact that singletons are closed sets. The theoretical domain, instead, corresponds to the standard Borel algebra.

Similarly, one can imagine verifiable statements for relationships (i.e. \statement{when the temperature of the mercury column is between 24 and 25 C, its height is between 24 and 25 mm}) and for statistical variables (i.e. \statement{if this coin is tossed enough times, the ratio of heads will be between 45\% and 55\%}). Ultimately the general theory will need to show precisely what mathematical structures map to experimental domains formed by statements of these types and under what assumptions.

@@

The general theory presented in this paper can be briefly summarized as follows. Science deals with statements that can be shown to be true experimentally. A scientific domain (i.e. an experimental domain) will be defined by a set of verifiable statements. From these we construct the set of all possible predictions (i.e. the theoretical domain), and concentrate on those that fully predict all our measurements, the possibilities. By the very nature and properties of these objects, a topology and a $\sigma$-algebra is induced on the possibilities that retains the connection to which statements are verifiable and which statements constitute a valid prediction.

The framework will be further expanded by studying how additional assumptions constrain the verifiable statements. Assuming a quantity can be measured with arbitrary precision will lead to the real numbers with the standard topology. Similarly, other assumptions will lead to other mathematical structures. This expansion is the aim of future work.

One benefit of the general theory is to bring to the forefront our hidden assumptions. If real numbers are associated with arbitrary precision, what happens in the case of limited precision, when we cannot assume precision can always be increased indefinitely? Would we still have real numbers? Would we be justified to use smooth manifolds or would we need an entirely different topological structure?

The other benefit is that no interpretation is required. All mathematical definitions are capturing well defined physical concepts. Therefore every theorem, every step in every proof will also have a well defined physical meaning.

We feel that this type of work is of significance for the scientific community in general and for the physics community in particular if we are ever to truly understand what scientific theories are meant to be describing.

>>>>

\newif\ifarchive
\archivetrue
%\archivefalse

\ifarchive
	\documentclass[letterpaper]{article}
\else
	\documentclass[smallextended]{svjour3}
\fi

%% Language and font encodings
%\usepackage[english]{babel}
%\usepackage[utf8x]{inputenc}
%\usepackage[T1]{fontenc}

%% Sets page size and margins
%\usepackage[letterpaper,top=1in,bottom=1in,left=1in,right=1in,marginparwidth=0.5in]{geometry}

%\usepackage{lineno,hyperref}
\usepackage{tikz}

\ifarchive
% Theorem definitions using amsthm
\usepackage{amsmath, amsthm, amsfonts}

@@

\begin{abstract}
	We lay the groundwork for a formal framework that studies scientific theories and can serve as a unified foundation for the different theories within physics. We define a scientific theory as a set of verifiable statements, assertions that can be shown to be true with an experimental test in finite time. By studying the algebra of such objects, we show that verifiability already provides severe constraints. In particular, it requires that a set of physically distinguishable cases is naturally equipped with the mathematical structures (i.e. second-countable Kolmogorov topologies and $\sigma$-algebras) that form the foundation of manifold theory, differential geometry, measure theory, probability theory and all the major branches of mathematics currently used in physics. This gives a clear physical meaning to those mathematical structures and provides a strong justification for their use in science. Most importantly it provides a formal framework to incorporate additional assumptions and constrain the search space for new physical theories.
\end{abstract}

@@

When considering physics as a discipline, one cannot help but notice that it is essentially composed of different theories and models loosely connected to each other, each with its own starting points and realm of applicability. Classical mechanics, electromagnetism, general relativity, quantum mechanics, thermodynamics and so on may share some general ideas but, in the end, they all have their own separate foundation, either in a different set of laws and principles or simply in positing a unique mathematical structure. This state of affairs is so entrenched in our field it somehow feels like the proper, if not the only, approach. But is it? Or does this approach actually hinder progress and understanding?

We can gain perspective by comparing to different fields of science and engineering. For example, in computer science theory one starts by defining symbolic languages and computational devices~\cite{Turing} and then shows that no algorithm exists that can correctly decide whether an arbitrary program terminates given an arbitrary input~\cite{Sipser}. In communication theory one defines channels and the information they carry~\cite{Shannon} and proves general results such as calculating the maximum rate at which information can be transmitted over a channel given a set amount of noise~\cite{Pierce}. In control theory one defines a system in terms of state, inputs and outputs~\cite{Brogan} and then looks for general strategies for control such as Kalman filtering~\cite{Kalman}. These theories, with their respective mathematical structures and results, are \emph{general} in the sense that they apply to all control systems, all communication systems and all computational devices. More specific topics in each field are constructed by further constraining those general mathematical structures. In other words, the subject matter has been properly defined formally and therefore the results follow simply from the mere definition of the problem.

In physics there is no equivalent. Classical mechanics, thermodynamics and quantum mechanics are not specific topics within some more general formal theory. But how would that work? As in the other general theories, we need to clarify what our starting points are and let the logic follow. The first requirement of a physical theory, or rather of any scientific theory in general, is that it is experimentally testable. So we would start by characterizing the properties and limitations of experimental verification. The second requirement, which is where physics starts, is identifying states and processes. After those are characterized in general, we can then specialize them for different cases. Some processes will be deterministic and reversible. Some processes will be non-deterministic and some will have equilibria. Some states can be broken into parts and some cannot. By adding different assumptions on states and processes we would recover classical Hamiltonian mechanics in one case, quantum mechanics in another and thermodynamics in yet another. In all these theories, some basic properties will be common simply because states need to be identified experimentally, and some properties will be different because the type of state or the type of process is different. This is what we mean by a general mathematical theory of experimental science.

%[rectangle, draw, fill=black!10, inner sep=5pt, text width=4cm, text badly centered, minimum height=1.2cm]
%every node/.style={draw, align = center},

\begin{figure}\label{twoQuestions}
	\centering
	\begin{tikzpicture}
	[child anchor = north, every node/.style={draw, align = center, inner sep=5pt},
	edge from parent/.style={draw, ->, -latex, stealth-},
	level 1/.style={level distance=1cm, sibling distance=6cm},
	level 2/.style={level distance=1.3cm, sibling distance=3cm}, sloped]
	\node {Experimental verifiability} [text width=4cm]
	child {node {States and processes} [text width={}]
		child {
			node {Classical \\ mechanics}
		}
		child {
			node {Quantum \\ mechanics}
		}
		child {
			node {Thermodynamics}
		}
		child {
			node {...}
		}
	};
	\end{tikzpicture}
	\caption{Overall structure for a general mathematical theory of experimental science. The starting point is experimental verification. States and processes are particular types of experimentally testable objects. Different theories in physics describe particular states under particular processes.}
\end{figure}

Such a general theory would force us to clarify our assumptions, thus the realm of applicability of each theory. It would always put physics at the center of our discussion, as physical ideas are the starting point for our formal framework and not after-the-fact interpretations. It would give science sturdier mathematical grounds, as each mathematical symbol is given a precise meaning and non-physical objects are excluded by construction. It would foster connections between different fields of knowledge: nature does not care about such divisions. It would provide a framework in which to pose new questions and solve old ones. In other words: it would provide a better foundation for physics and for the rest of the sciences. Developing a general mathematical theory for experimental science could be compared to what happened in mathematics during the first half of the last century, when it was reorganized using logic and set theory as its foundation, which had profound repercussions in the field. Our preliminary work in both physics~\cite{Carc1} and math~\cite{Carc2} convinced us that such a goal is possible and within reach.

The aim of the present work is to lay down the beginning of a general mathematical theory of experimental science, limited to the part concerning experimental verification. We start by defining verifiable statements: assertions whose truth can be verified experimentally in finite time. We study the logic of these statements, which is different from the standard Boolean logic because of finite time termination. We define experimental domains as a collection of verifiable statements that, if given an indefinite amount of time, we can keep testing, forever refining our knowledge. From each experimental domain we construct its possibilities: the different cases we can experimentally distinguish. The main result is that \textbf{each set of experimentally distinguishable possibilities comes equipped with a natural second-countable Kolmogorov topology, where each open set corresponds to a verifiable statement, and a natural $\sigma$-algebra, where each Borel set corresponds to a theoretical statement which gives predictions for verifiable ones.} For example, \statement{the mass of the particle is more than $0.4$ and less than $0.6$ MeV} is a verifiable statement precisely because $(0.4, 0.6)$ is an open set, while \statement{the mass of the particle is exactly $0.5$ MeV} is not verifiable because $[0.5,0.5]$ is not an open set.\footnote{The standard topology of the real numbers keeps track of finite precision measurements.} From that general result we can derive the following conclusion: \textbf{any set of physically distinguishable cases has at most cardinality of the continuum}. That is, the only thing we can do to distinguish a particular case is run the tests for a countable set of statements, the output of which can be imagined as a countable sequence of true and false. This is equivalent to the binary representation of a real number. In a nutshell, \textbf{experimental verification by itself guarantees us the existence of two mathematical structures that are the foundations of most tools used in physics}, such as differential geometry (Riemannian and symplectic), measure theory, probability theory and many others.

%In this paper we briefly present the starting point for our general mathematical theory of experimental science. It is the product of a decade spent in reverse engineering the laws of particle mechanics~\cite{Carc1} and our subsequent work on formalizing the starting points~\cite{Carc2}. Our overall program is large and ambitious and a journal article cannot treat it comprehensively. Therefore we will only present an overview of the basic concepts, which are already well consolidated and provide a good template to understand our approach. We will only provide a sketch for the proofs, which can be found in the draft of our open access book~\cite{Carc3} together with more details, discussions and insights.

While this result is remarkable by itself, the most interesting aspect is that it can be used as a cornerstone for a more general theory. Conceptually, in the same way that topological spaces keep track of what can be verified experimentally, manifolds keep track of objects that can be identified by continuous quantities, differentiable manifolds keep track of objects that allow a density to be defined over them and symplectic manifolds (i.e. phase space of classical mechanics) allow the density to be coordinate invariant (i.e. observer independent). By doing all this work formally we will be forced to recognize the assumptions that go into those structures, and understand if and how they fail. For example, we are currently in the process of identifying the necessary and sufficient conditions such that a set of verifiable statements defines a continuous quantity. While, as one would expect, these can only correspond to an idealized case where one pretends a better precision measurement can always be achieved, they also tell us precisely how this idealized case fails which can be relevant to work on Planck scale physics. But whatever other tools we will have in that case, we \emph{know} we will always have at least a topology and a $\sigma$-algebra.

As another example, note that the set of discontinuous functions from $\mathbb{R}$ to $\mathbb{R}$ has cardinality greater than the continuum and therefore cannot represent a set of physically distinguishable objects. This means that the square integrable space of functions typically used in quantum mechanics is far too big, and, in fact, it contains unphysical elements, such as states with infinite energy. However, the Schwartz space, like the set of continuous functions, has cardinality of the continuum and can be given a second-countable Kolmogorov topology.\footnote{It also has other properties that map to physical requirements. For example, quantities of the form $\langle \psi|Q^n P^m|\psi\rangle$ are always finite and the Fourier transform is always well defined.} To our understanding, the reason why Hilbert spaces are mathematically useful is because one can take limits and use the vector space norm to converge. However, we may instead be able to use the corresponding $\sigma$-algebra to take limits, like one does in probability and measure theory. In other words, we may be able to construct, within this framework, a more mathematically sound and physically motivated foundation for quantum mechanics that includes the objects and only the objects that are physically meaningful.

As a last example, we also believe we can form a connection to computer science. We can characterize the output of a computational device as a set of verifiable statements (e.g. \statement{the third bit of output is 1}) that are a function of the input, which can also be characterized as verifiable statements (e.g. \statement{the second bit of the input is 0}). In this context, all possible inputs form the possibilities of our space, as they define all possible outputs of the computation. Now suppose we have a function from the input to a Boolean value. Then we can write the statements \statement{the input is such that the output of the function is true} and \statement{the input is such that the output of the function is false}. If both these statements correspond to open sets on the possible inputs then they are verifiable statements and we construct a test (i.e. a program) that will always terminate. That is, we can combine tools from topology and computer science to answer a question of the type: is this scientific problem testable or computable? In a similar way, connections to information theory can also be established: a deterministic and reversible process, in fact, is one for which description of the input is equivalent to the description of the output, which means information entropy is conserved. These links illustrate the potential as a theory of experimental science, and not just physics.

We hope that it is clear by now that a general mathematical theory of experimental science is missing, is possible and would be extremely useful. Whether you work on condensed matter, quantum thermodynamics, complex systems, particle physics, gravitation or grand unified theories, if what you are doing is science you will have, at some point, experimental verification. You will have problems (often the most interesting ones) that arise from questions such as: what is it that I can measure? How do I do it? Under what assumptions do the quantities in my theory map to those measurements? When are they even well defined? What can I compute within my theory? What is testable? Currently, you do not have a general framework to pose those questions. And it turns out you do not need a completely ad-hoc one for your specific field. The mere \emph{logic} of experimental verification requires a particular structure. And there is a lot more to experimental verification than its logic: the quantity itself has to, at least in a sense, exist over a finite period of time to be measurable, there must exist a process to transfer that information to a measurement device, the system cannot be assumed to be always completely isolated as it would not be physically accessible. If you are working on foundational aspects, chances are you may already be thinking about some of these issues as they relate to your field. But what parts of those problems are specific to your field? How much can be treated generally, so that not only each field has less work to do but also we have a common language across fields? What are the assumptions specific to your field and those that are more general? A general mathematical theory of experimental science would provide you with a formal framework to precisely pose and possibly answer those questions. As the starting points are clarified, the foundations of different fields may be affected, or even partially integrated, and potentially change how they are taught as well.

The irony is that while the project is extremely ambitious, the biggest obstacles are not, at this point, technical. The biggest obstacles are sociological. On one side the nature of this work is extremely interdisciplinary. On the other side math and physics are divided into fields and sub-fields that are increasingly narrow.\footnote{We attended a topology conference where there seemed to be a disconnect between what mathematicians and condensed matter physicists meant by topology, to the point that they frequently talked past each other.} This means that this work does not fit naturally within any community while requiring technical expertise from many. To obviate these problems, borrowing practices from the open source software community, we are developing our main body of work in the open~\cite{Carc3}, so that we can address the topic in its entirety, and our colleagues in philosophy, mathematics and physics can all check their respective parts. This article is the product of that process and focuses on current results of greater interest to physicists, omitting more mathematical and philosophical details that can be found in the broader work.

@@

As the starting point of our general theory, we introduce the following guiding principle:

@@

This means that science concerns itself only to the study of assertions that have a well defined truth value which can be verified experimentally. The issue at hand is to formally capture this informal intuition. We begin with the following common definition.

@@

Next we need to define our truth bearer. In mathematical logic, this is typically a well formed formula. This cannot work for us: what we are interested in is the meaning of the assertion and not how it is expressed. For example, \statement{this animal is a dog} and \statement{questo animale \`e un cane} represent the same fact expressed in different languages. As science is universal, it should not matter the language, units or reference system used to make an assertion.\footnote{In the same vein, a statement is not necessarily one sentence as it has to include all the information needed to understand what exactly we are taking about. These are the type of philosophical issues that are discussed at greater length in~\cite{Carc3}.} While we will still use standard mathematics for the formal system, we introduce a variation of algebraic logic to represent our ``informal'' assertions. We start by defining our truth bearer as:

@@

\section{Brief discussion and examples}

The connection we outlined creates a strong bridge between the standard mathematical structures used in physics and their meaning. Every theorem, every proof on those structures can now be given a direct physical meaning as well. And given the foundational nature of topologies and $\sigma$-algebras, we can imagine extending this framework to measure theory, differential geometry, symplectic geometry, Riemannian geometry, probability theory and so on.

To give an example of how it works, consider an experimental domain where the basis is composed of statements like \statement{this quantity is more than $x$ but less than $y$} where $x$ and $y$ are different values that can be arbitrarily close. For example, \statement{the distance between the earth and the moon is more than 384 but less than 385 thousand Km}. Each verifiable statement of the basis corresponds to an open interval of the real line, therefore we find a correspondence between arbitrary precision measurements and the standard topology on the reals, as this is the one generated by open intervals. Note, in fact, that this topology is second-countable and at least Kolmogorov.

The possibilities (e.g. \statement{the mass of the photon is precisely zero}) are not themselves verifiable since infinite precision measurements of a continuous quantity are not possible. Yet, their negation (i.e. \statement{the mass of the photon is not precisely zero}) could be verified in practice. This is expressed mathematically by the fact that singletons are closed sets. The theoretical domain, instead, corresponds to the standard Borel algebra, which includes closed and half-open sets but not all possible sets of integers, many of which cannot be characterized by a well formed formula.

Similarly, one can imagine verifiable statements for relationships (e.g. \statement{when the temperature of the mercury column is between 24 and 25 C, its height is between 24 and 25 mm}) and for statistical variables (e.g. \statement{if this coin is tossed enough times, the fraction of heads will be between 45\% and 55\%}). Ultimately the general theory will need to show precisely what mathematical structures map to experimental domains formed by statements of these types and under what assumptions. But the idea is that, once you have specified the verifiable statements and their logical relationships, you have already specified the experimental domain and nothing else needs to be added. In other words: \textbf{the points of the space and their mathematical structures are fully specified by what can be measured within the theory}.

@@

Science is based on experimental verification. Experimental verification has its own logic. This logic imposes a mathematical structure which we characterized in terms of experimental domains (collections of verifiable statements), theoretical domains (collections of predictions for verifiable statements) and possibilities (the cases that can be distinguished experimentally). This mathematical structure leads naturally to topological spaces and $\sigma$-algebras, the foundation of many of the tools used in physics and science in general. Not only does this result clarify why those tools are so successful in science, it provides a direct physical meaning to those structures and a solid foundation upon which to create a general mathematical theory of experimental science.

In this light, what is most remarkable about this work is not its results, but that \emph{it can be done in the first place}. That there is a way to construct physical theories that forces us to spell out our physical assumptions, and that clarifies in a rigorous way what result is a consequence of what assumption. We can therefore analyze and compare new starting points with the discipline and thoroughness of modern mathematics instead of just rummaging through the bag of mathematical tools. But to do that, we have to create a space in the scientific community that is conducive to this kind of broad interdisciplinary work. We need generalists that can recognize how a detail in measure theory relates to information entropy or Hamiltonian mechanics. We need to take the mathematical structures created by mathematicians for their needs, break them apart and recombine them to find mathematical structures most suited and meaningful for science. Scientific knowledge has expanded quite considerably in the last century, maybe it is time for a moment of synthesis and consolidation.

We believe the development of a general mathematical theory of experimental science would be an important accomplishment for the scientific community and that it would be beneficial in ways we cannot yet even imagine. All we know is that a clearer understanding of our starting points cannot but help to provide insights into the current theories and suggest strategies to address long standing problems.
